---
output: 
  html_document
self_contained: yes
runtime: shiny
---

# Evaluating combined effects of restoration {.tabset}

```{r globals, echo = F, message = F, warning = F}
knitr::opts_chunk$set(message = F, warning = F, echo = F)

library(tidyverse)
library(ggmap)
library(lubridate)
library(geosphere)
library(stringi)
library(tibble)
library(scales)
library(shiny)
library(sf)
library(sp)
library(gridExtra)

data(restdat)
data(reststat)
data(wqdat)
data(wqstat)
data(map)
data(grdsres)

# source R files
source('R/get_chg.R')
source('R/get_clo.R')
source('R/get_dat.R')
source('R/get_lik.R')

# base map
# ext <- make_bbox(wqstat$lon, wqstat$lat, f = 0.2)
# get_stamenmap(ext, zoom = 11, maptype = "toner-lite", where = getwd())
pbase <- ggmap(map) +
  theme_bw(base_family = 'serif', base_size = 16) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

```{r inpts}
column(4, 
       selectInput('sta', 'Select water quality station:', choices = sort(wqstat$stat), selected = 55)
       )
```

```{r rctvs}
# filtered wqdat by stat
wqdatsta <- reactive({
  
  # inputs
  sta <- input$sta
  
  wqdat %>% 
    filter(stat %in% sta)
  
})

# filtered wqstat by station
wqstatsta <- reactive({
  
  # inputs
  sta <- input$sta
  
  wqstat %>% 
    filter(stat %in% sta)
  
})

```

```{r}
renderPlot({
  
  # inputs
  wqdatsta <- wqdatsta()
  wqstatsta <- wqstatsta()
  
  # wq ts
  p2 <- ggplot(wqdatsta, aes(x = datetime, y = chla)) + 
    geom_line() +
    geom_point() +
    theme_bw(base_family = 'serif') +
    theme(
      axis.title.x = element_blank()
    ) +
    scale_y_log10('log10-Chla')
  
  # map
  p1 <- pbase + 
    geom_point(data = wqstatsta, aes(x = lon, y = lat), colour = 'tomato1', size = 6) +
    geom_text(data = wqstat, aes(x = lon, y = lat, label = stat), size = 3)

  grid.arrange(p1, p2, widths = c(0.3, 0.7), ncol = 2)
  
}, height = 300, width = 900)
```

```{r inptsmp}
# wq year slice window
column(4,
       numericInput("yrdf", "Choose year diff:", 5)
       ) 
    
# wq site match upper limit
column(4, 
    numericInput("mtch", "Choose restoration matches:", 5)
)

# jitr
column(4, 
    sliderInput("jitr", "Jitter overlap restorations:", min = 0, max = 50, value = 0)
)
```

```{r rctvsmp}
# wq stats matched to rest stats
wqmtch <- reactive({
  
  # reactive inputs
  mtch <- input$mtch
  wqstatsta <- wqstatsta()

  # get wqmtch
  out <- get_clo(restdat, reststat, wqstatsta, resgrp = 'type', mtch = mtch)

  return(out)
  
})

# yr windows with matched rest sites
yrwin <- reactive({
  
  # input
  wqstatsta <- wqstatsta()
  wqdatsta <- wqdatsta()
  wqmtch <- wqmtch()
  yrdf <- input$yrdf
  
  # get temp matches and avgs
  yrwin <- get_chg(wqdatsta, wqmtch, wqstatsta, restdat, wqvar = 'chla', yrdf = yrdf, chgout = T)
  
  return(yrwin)
  
})
```

# {.tabset}

## (1) Map

The water quality station is matched to the closest `r renderText({input$mtch})` projects for each project type.

```{r}
renderPlot({
  
  # inputs
  wqmtch <- wqmtch()
  wqstatsta <- wqstatsta()
  jitr <- input$jitr
  
  # combine lat/lon for the plot
  toplo <- wqmtch %>%
    left_join(wqstatsta, by = 'stat') %>%
    left_join(reststat, by = 'id') %>%
    rename(
      `Restoration type` = resgrp,
      `Distance (dd)` = dist
    ) %>% 
    mutate(
      lat.y = ifelse(duplicated(lat.y), jitter(lat.y, factor = jitr), lat.y),
      lon.y = ifelse(duplicated(lon.y), jitter(lon.y, factor = jitr), lon.y)
      )
      
  # restoration project grouping column
  restall <- left_join(restdat, reststat, by = 'id')
  names(restall)[names(restall) %in% 'type'] <- 'Restoration type'
  
  # rngs
  xrng <- range(c(toplo$lon.x, toplo$lon.y))
  yrng <- range(c(toplo$lat.x, toplo$lat.y))
  
  # base extension
  p <- pbase +
    geom_point(data = restall, aes(x = lon, y = lat), size = 1, pch = 21) +
    geom_segment(data = toplo, aes(x = lon.x, y = lat.x, xend = lon.y, yend = lat.y, alpha = -`Distance (dd)`, linetype = `Restoration type`), size = 1) +
    geom_point(data = toplo, aes(x = lon.y, y = lat.y, fill = `Restoration type`), size = 3, pch = 21) +
    geom_point(data = wqstatsta, aes(x = lon, y = lat), size = 5, colour = 'tomato1') +
    scale_alpha(guide = F) 
  
  # whole map
  p1 <- p +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(), 
      legend.position = 'none'
    )

  
  # sub map
  p2 <- p +
    scale_y_continuous(limits = yrng) +
    scale_x_continuous(limits = xrng) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(), 
      legend.position = 'top', 
      legend.title = element_blank()
    )

  grid.arrange(p1, p2, ncol = 2)

}, width = 900, height = 650)
```

## (2) Time slices

For each of `r renderText({input$mtch})` restoration project types, the water quality observations that occurred `r renderText({input$yrdf})` years before and after the completion of the project were selected. The date of completion for each project is shown with the thicker vertical line. 

```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # wq ts
  p <- ggplot() + 
    geom_rect(data = yrwin, aes(xmin = dtend, xmax = dtaft, ymin = 0, ymax = Inf, group = resgrp, fill = resgrp), alpha = 0.7) +
    geom_point(data = wqdatsta, aes(x = datetime, y = chla)) +
    geom_vline(data = yrwin, aes(xintercept = date, colour = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    facet_grid(rnk ~ resgrp) + 
    scale_y_log10('log10-Chla')
  
  return(p)
  
}, width = 900, height = 600)  
```

## (3) Before/after avg.

The water quality observations for the `r renderText({input$yrdf})` years before and after the selected slices were then averaged. 
```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # to plot
  toplo <- yrwin %>% 
    select(rnk, resgrp, wts, bef, aft) %>% 
    gather('var', 'val', bef, aft) %>% 
    mutate(var = factor(var, levels = c('bef', 'aft')))
  
  # wq chgs bar plot
  p <- ggplot() + 
    geom_bar(data = toplo, stat = 'identity', aes(x = var, y = val, fill = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    facet_grid(rnk ~ resgrp) + 
    scale_y_log10('log10-Chla')
  
  return(p)
  
}, width = 900, height = 600)  
```

## (4) Before/after avg. difference

The before and after averages were then differenced.

```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # to plot
  toplo <- yrwin %>% 
    select(rnk, resgrp, wts, bef, aft) %>% 
    mutate(
      difv = aft - bef,
      xvr = 'Diff'
      )
  
  # wq chgs bar plot
  p <- ggplot() + 
    geom_bar(data = toplo, stat = 'identity', aes(x = xvr, y = difv, fill = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    facet_grid(rnk ~ resgrp) + 
    ylab('Chla difference') +
    geom_hline(yintercept = 0, linetype = 'dotted')
  
  return(p)
  
}, width = 900, height = 600)  
```

## (5) Average of averages

Then the averaged differences are averaged across the `r renderText({input$mtch})` projects per project type.

```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # to plot
  toplo <- yrwin %>% 
    select(rnk, resgrp, wts, bef, aft) %>% 
    mutate(
      difv = aft - bef,
      xvr = 'Diff'
      ) %>% 
    group_by(resgrp) %>% 
    summarise(
      avedf = mean(difv, na.rm = T)
    )

  # wq chgs bar plot
  p <- ggplot() + 
    geom_bar(data = toplo, stat = 'identity', aes(x = resgrp, y = avedf, fill = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    ylab('Chla difference, averages') +
    geom_hline(yintercept = 0, linetype = 'dotted')
  
  return(p)
  
}, width = 900, height = 250)  
```