---
output: 
  html_document
self_contained: yes
runtime: shiny
---

# Evaluating combined effects of restoration {.tabset}

```{r globals, echo = F, message = F, warning = F}
knitr::opts_chunk$set(message = F, warning = F, echo = F)

library(tidyverse)
library(ggmap)
library(lubridate)
library(geosphere)
library(stringi)
library(tibble)
library(scales)
library(shiny)
library(sf)
library(sp)
library(gridExtra)
library(multcompView)

data(restdat)
data(reststat)
data(wqdat)
data(wqstat)
data(map)
data(grdave)

# source R files
source('R/get_chg.R')
source('R/get_clo.R')
source('R/get_dat.R')
source('R/get_lik.R')

# base map
# ext <- make_bbox(wqstat$lon, wqstat$lat, f = 0.2)
# get_stamenmap(ext, zoom = 11, maptype = "toner-lite", where = getwd())
pbase <- ggmap(map) +
  theme_bw(base_family = 'serif', base_size = 16) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

# Individual stations {.tabset}

```{r inpts}
column(4, 
       selectInput('sta', 'Select water quality station:', choices = sort(wqstat$stat), selected = 55)
       )
```

```{r rctvs}
# filtered wqdat by stat
wqdatsta <- reactive({
  
  # inputs
  sta <- input$sta
  
  wqdat %>% 
    filter(stat %in% sta)
  
})

# filtered wqstat by station
wqstatsta <- reactive({
  
  # inputs
  sta <- input$sta
  
  wqstat %>% 
    filter(stat %in% sta)
  
})

```

```{r}
renderPlot({
  
  # inputs
  wqdatsta <- wqdatsta()
  wqstatsta <- wqstatsta()
  
  # wq ts
  p2 <- ggplot(wqdatsta, aes(x = datetime, y = chla)) + 
    geom_line() +
    geom_point() +
    theme_bw(base_family = 'serif') +
    theme(
      axis.title.x = element_blank()
    ) +
    scale_y_log10('log10-Chla')
  
  # map
  p1 <- pbase + 
    geom_point(data = wqstatsta, aes(x = lon, y = lat), colour = 'tomato1', size = 6) +
    geom_text(data = wqstat, aes(x = lon, y = lat, label = stat), size = 3)

  grid.arrange(p1, p2, widths = c(0.3, 0.7), ncol = 2)
  
}, height = 300, width = 900)
```

```{r inptsmp}
# wq year slice window
column(4,
       numericInput("yrdf", "Choose year diff:", 5)
       ) 
    
# wq site match upper limit
column(4, 
    numericInput("mtch", "Choose restoration matches:", 5)
)

# jitr
column(4, 
    sliderInput("jitr", "Jitter overlap restorations:", min = 0, max = 50, value = 0)
)
```

```{r rctvsmp}
# wq stats matched to rest stats
wqmtch <- reactive({
  
  # reactive inputs
  mtch <- input$mtch
  wqstatsta <- wqstatsta()

  # get wqmtch
  out <- get_clo(restdat, reststat, wqstatsta, resgrp = 'type', mtch = mtch)

  return(out)
  
})

# yr windows with matched rest sites
yrwin <- reactive({
  
  # input
  wqstatsta <- wqstatsta()
  wqdatsta <- wqdatsta()
  wqmtch <- wqmtch()
  yrdf <- input$yrdf
  
  # get temp matches and avgs
  yrwin <- get_chg(wqdatsta, wqmtch, wqstatsta, restdat, wqvar = 'chla', yrdf = yrdf, chgout = T)
  
  return(yrwin)
  
})
```

## (1) Map

The water quality station is matched to the closest `r renderText({input$mtch})` projects for each project type.

```{r}
renderPlot({
  
  # inputs
  wqmtch <- wqmtch()
  wqstatsta <- wqstatsta()
  jitr <- input$jitr
  
  # combine lat/lon for the plot
  toplo <- wqmtch %>%
    left_join(wqstatsta, by = 'stat') %>%
    left_join(reststat, by = 'id') %>%
    rename(
      `Restoration type` = resgrp,
      `Distance (dd)` = dist
    ) %>% 
    mutate(
      lat.y = ifelse(duplicated(lat.y), jitter(lat.y, factor = jitr), lat.y),
      lon.y = ifelse(duplicated(lon.y), jitter(lon.y, factor = jitr), lon.y)
      )
      
  # restoration project grouping column
  restall <- left_join(restdat, reststat, by = 'id')
  names(restall)[names(restall) %in% 'type'] <- 'Restoration type'
  
  # rngs
  xrng <- range(c(toplo$lon.x, toplo$lon.y))
  yrng <- range(c(toplo$lat.x, toplo$lat.y))
  
  # base extension
  p <- pbase +
    geom_point(data = restall, aes(x = lon, y = lat), size = 1, pch = 21) +
    geom_segment(data = toplo, aes(x = lon.x, y = lat.x, xend = lon.y, yend = lat.y, alpha = -`Distance (dd)`, linetype = `Restoration type`), size = 1) +
    geom_point(data = toplo, aes(x = lon.y, y = lat.y, fill = `Restoration type`), size = 3, pch = 21) +
    geom_point(data = wqstatsta, aes(x = lon, y = lat), size = 5, colour = 'tomato1') +
    scale_alpha(guide = F) 
  
  # whole map
  p1 <- p +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(), 
      legend.position = 'none'
    )

  
  # sub map
  p2 <- p +
    scale_y_continuous(limits = yrng) +
    scale_x_continuous(limits = xrng) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(), 
      legend.position = 'top', 
      legend.title = element_blank()
    )

  grid.arrange(p1, p2, ncol = 2)

}, width = 900, height = 650)
```

## (2) Time slices

For each of `r renderText({input$mtch})` restoration project types, the water quality observations that occurred `r renderText({input$yrdf})` years before and after the completion of the project were selected. The date of completion for each project is shown with the thicker vertical line. 

```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # wq ts
  p <- ggplot() + 
    geom_rect(data = yrwin, aes(xmin = dtend, xmax = dtaft, ymin = 0, ymax = Inf, group = resgrp, fill = resgrp), alpha = 0.7) +
    geom_point(data = wqdatsta, aes(x = datetime, y = chla)) +
    geom_vline(data = yrwin, aes(xintercept = date, colour = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    facet_grid(rnk ~ resgrp) + 
    scale_y_log10('log10-Chla')
  
  return(p)
  
}, width = 900, height = 600)  
```

## (3) Before/after avg.

The water quality observations for the `r renderText({input$yrdf})` years before and after the selected slices were then averaged. 
```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # to plot
  toplo <- yrwin %>% 
    select(rnk, resgrp, wts, bef, aft) %>% 
    gather('var', 'val', bef, aft) %>% 
    mutate(var = factor(var, levels = c('bef', 'aft')))
  
  # wq chgs bar plot
  p <- ggplot() + 
    geom_bar(data = toplo, stat = 'identity', aes(x = var, y = val, fill = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    facet_grid(rnk ~ resgrp) + 
    scale_y_log10('log10-Chla')
  
  return(p)
  
}, width = 900, height = 600)  
```

## (4) Before/after avg. difference

The before and after averages were then differenced.

```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # to plot
  toplo <- yrwin %>% 
    select(rnk, resgrp, wts, bef, aft) %>% 
    mutate(
      difv = aft - bef,
      xvr = 'Diff'
      )
  
  # wq chgs bar plot
  p <- ggplot() + 
    geom_bar(data = toplo, stat = 'identity', aes(x = xvr, y = difv, fill = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    facet_grid(rnk ~ resgrp) + 
    ylab('Chla difference') +
    geom_hline(yintercept = 0, linetype = 'dotted')
  
  return(p)
  
}, width = 900, height = 600)  
```

## (5) Average of averages

Then the averaged differences are averaged across the `r renderText({input$mtch})` projects per project type.

```{r}
renderPlot({
  
  # input
  yrwin <- yrwin()
  wqdatsta <- wqdatsta()

  # to plot
  toplo <- yrwin %>% 
    select(rnk, resgrp, wts, bef, aft) %>% 
    mutate(
      difv = aft - bef,
      xvr = 'Diff'
      ) %>% 
    group_by(resgrp) %>% 
    summarise(
      avedf = mean(difv, na.rm = T)
    )

  # wq chgs bar plot
  p <- ggplot() + 
    geom_bar(data = toplo, stat = 'identity', aes(x = resgrp, y = avedf, fill = resgrp)) +
    theme_bw(base_family = 'serif', base_size = 20) +
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'none', 
      panel.border = element_rect(fill = NA),
      strip.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.spacing = grid::unit(0, "lines")
    ) +
    ylab('Chla difference, averages') +
    geom_hline(yintercept = 0, linetype = 'dotted')
  
  return(p)
  
}, width = 900, height = 250)  
```

# Evaluation of all differences

This map shows a summary of all average difference values in Chlorophyll for selected year window and number of restoration project matches.  The restoration project grouping (simple: top, complex: type) can also be toggled.  The points are displayed by size to show magnitude of the change and color to show direction of the change (red is increase, green is decrease). These are the same values that are shown in tab 5 but mapped to all stations for any variable combination from the selection tabs.

```{r}
data(grdave)

column(12, 
       
       column(4, 
              selectInput('yrdfgrd', 'Select year window:', choices = unique(grdave$yrdf))
              ),
       
       column(4, 
              selectInput('mtchgrd', 'Select restoration site matches:', choices = unique(grdave$mtch))
              ),
       
       column(4,  
              selectInput('resgrpgrd', 'Select restoration group level:', choices = unique(grdave$resgrp))
              )
       
)

# filter results      
datdf <- reactive({
  
  # input
  mtchgrd <- input$mtchgrd
  yrdfgrd <- input$yrdfgrd
  resgrpgrd <- input$resgrpgrd

  out <- grdave %>% 
    filter(yrdf %in% yrdfgrd) %>% 
    filter(mtch %in% mtchgrd) %>% 
    filter(resgrp %in% resgrpgrd) %>% 
    pull(value) %>% 
    .[[1]]
  
  return(out)
  
})

# get multiple comparison letters and overall pvalue
letvl <- reactive({
  
  # input
  datdf <- datdf()
  
  # anova and HSD comparison letters
  mod <- aov(avedf ~ resgrp, data = datdf)
  modhsd <- TukeyHSD(mod) %>% 
    .$resgrp %>% 
    data.frame
  pvals <- modhsd$p.adj
  names(pvals) <- rownames(modhsd)
  lets <- multcompLetters(pvals)
  if('Letters' %in% names(lets))
    lets <- lets$Letters
  lets <- lets %>% 
    enframe('resgrp', 'lets')
  
  # get average and max chl chng for color mapping
  datdf <- datdf %>% 
    group_by(resgrp) %>% 
    mutate(
      maxdf = max(avedf, na.rm = T)
      ) %>% 
    ungroup
  
  # join lets with max dif values for labels
  lets <- datdf %>% 
    select(resgrp, maxdf) %>% 
    unique %>% 
    mutate(resgrp = as.character(resgrp)) %>% 
    left_join(lets, by = 'resgrp')
  
  # overal pvalue
  pval <- mod %>% 
    summary %>% 
    .[[1]] %>% 
    .[["Pr(>F)"]] %>% 
    .[1] %>%
    round(2)
  
  out <- list(
    lets = lets, 
    pval = pval
  )
  
  return(out)
  
})

# test if df values diff from zero, by group
difzr <- reactive({

  # input
  datdf <- datdf()
  
  tst <- datdf %>% 
    group_by(resgrp) %>% 
    nest %>% 
    mutate(
      tst = map(data, ~ t.test(x = .$avedf)),
      pvl = map(tst, function(x){
        
        ifelse(x$p.value > 0.05, 'mean = 0', 
          ifelse(x$statistic > 0, 'mean > 0', 'mean < 0'))
       
        })
    ) %>% 
    select(resgrp, pvl) %>% 
    unnest
  
  return(tst)

})
```

```{r}
renderPlot({
  
  # input
  datdf <- datdf()

  datdfeval <- datdf %>% 
    left_join(wqstat, by = 'stat') %>% 
    rename(`Chlorophyll change` = avedf)

  # map
  p1 <- ggplot() + 
    geom_point(data = datdfeval, aes(x = lon, y = lat, fill = `Chlorophyll change`, size = abs(`Chlorophyll change`)), pch = 21, colour = 'black') +
    scale_fill_gradient2(low = 'lightgreen', mid = 'white',  high = 'tomato1', midpoint = 0) + 
    facet_grid(~ resgrp, switch = 'x') +
    coord_map() +
    theme_void(base_family = 'serif', base_size = 18) +
    theme(
      legend.position = 'top',
      strip.placement = 'outside'
    ) +
    scale_size(range = c(1, 11), guide = F) +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10))
  
  return(p1)
  
}, height = 300)

```

This plot shows overall differences in chlorophyll changes as a function of restoration project types.  The model is evaluated based on the aggregated differences across all sites given the selected number of matches and year windows.  The overall p-value is the ANOVA result testing for variation in chlorophyll differences by one or more project types and the letters in the plot show groups that are significantly different based on pairwise comparisons using adjust p-values.  The additional text shows whether the mean chlorophyll differences within each group are significantly different from zero.  

```{r}
renderPlot({
  
  # input
  datdf <- datdf()
  difzr <- difzr()
  letvl <- letvl()
  mtchgrd <- input$mtchgrd
  yrdfgrd <- input$yrdfgrd

  # get average chl chng for color mapping
  datdf <- datdf %>% 
    group_by(resgrp) %>% 
    mutate(
      grpdf = mean(avedf, na.rm = T)
      ) %>% 
    ungroup
  
  # multiplecomparison letters and diff values from difzr
  lets <- letvl$lets %>% 
    left_join(difzr, by = 'resgrp') %>% 
    unite(lets, c('lets', 'pvl'), sep = ', ')

  # title 
  titl <- letvl$pval %>% 
    paste0('ANOVA model pval = ', ., ' based on ', mtchgrd, ' projects matched to each station, +/- ', yrdfgrd, ' years')
  
  # map
  p1 <- ggplot() + 
    geom_hline(yintercept = 0, linetype = 'dashed', size = 1) +
    geom_boxplot(data = datdf, aes(x = resgrp, y = avedf, fill = grpdf)) +
    scale_fill_gradient2('Within-group average', low = 'lightgreen', mid = 'white',  high = 'tomato1', midpoint = 0) + 
    geom_text(data = lets, aes(x = resgrp, y = maxdf, label = lets), hjust = -0.1, size = 6) +
    theme_bw(base_family = 'serif', base_size = 18) +
    theme(
      legend.position = 'top'
    ) +
    scale_y_continuous('Chlophyll change') +
    scale_x_discrete('Restoration group') +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10)) +
    ggtitle(subtitle = titl, label = '')
  
  return(p1)
  
}, height = 450)
```

